Untwisting the Boerdijk-Coxeter Helix

### 2017-05-27 doc/TetrahelixUntwist.tex
formatting

The vertices of the tetrahedra lie upon
three helices about the central axis.
The \\
Tetrobot\cite{readglussbot,TetrobotBook} uses the regularity of
this geometry to make a tentacle-like robot that can crawl like a slug
or mollusc.  These modular robotic systems use mechanical actuators which can change their length, connected by special joints,
such as the 3D printable Song-Kwon-Kim\cite{song2003spherical} joint or the CMS joint\cite{HamlinSandersonCMS} used in the original Tetrobot,
which allow many members to meet in a single point.
Such machines can
follow purely regular mathematical models such as the Boerdijk-Coxeter
helix or the Octet Truss\cite{richard1961synergetic}.

@@

Ideally we would have a simple formula for defining the nodes based on
any curvature or pitch we choose. It is a goal of
this paper to relate the Cartesian coordinate approach and the member-length approach to
generating a tetrahelix continuum.

@@

R.W. Gray's website\cite{graytetrahelix}, repeating a formula by Coxeter\cite{coxeter1985simplicial} in a more accessible form, gives the Cartesian coordinates $\begin{bmatrix}
           x \\
           y \\
           z
         \end{bmatrix}$
for a counter-clockwise BC Helix in a right-handed coordinate system:
\begin{equation}
  \label{graycoxeter}
\vec{V}(n) =
\left [
  \begin{tabular}{c}
   $ r_{bc} \cos{n \theta_{bc}} $\\
   $ r_{bc} \sin{n \theta_{bc}} $\\
   $ n h_{bc}  $
  \end{tabular}
  \right ],
\text{where:}
  \begin{tabular}{c}
 $ r_{bc} = \frac{3\sqrt{3}}{10} \approx 0.5196 $\\
 $ h_{bc} = 1/\sqrt{10} \approx 0.3162 $ \\
 $ \theta_{bc} = \arccos(-2/3) $  \text{ ,}\\
  \end{tabular}      
\end{equation}
and where $n$ represents each integer numbered node in succession on every colored rail.

The apparent rotation of a vertex on an outer-edge, that is $\vec{V}(n)$ relative from $\vec{V}(n+3)$
for any integer $n$
in \eqref{graycoxeter}, is $3 \theta_{bc} - 2\pi$.

This formula defines a helix, but it is not any of the apparent helices, or \emph{rail} helices, of the
BC helix, but rather one that winds much more rapidly through all
nodes. To a designer of tetrahelices, it is more natural to think of
the three helices which are visually apparent, that is, those three
which are closely approximated by the outer edges or rails of
the BC helix. We think of each of these three rails as being a different color: red, blue, or yellow.
This situation is illustrated in \cref{fig:helixcomparison}, wherein the black helix represents that
generated by \cref{graycoxeter},
and the colored helices are generated by \cref{eq:colored}.

In order to develop the continuum of slightly irregular tetrahelices described in \cref{sec:continuum},
we need a formula that gives us the nodes of just
one rail helix, denoted by color $c$ and integer node number $n$:
\[
(\forall n \in \mathbb{Z}, \forall c \in \{0,1,2\} : \vec{H}_{BCcolored}(n,c) = \vec{V}(3n +c)) .
\]

Such a helix can be written:
\begin{equation}
  \label{eq:colored}
 \vec{H}_{BCcolored}(n,c) =
\left [
  \begin{tabular}{c}
   $ r_{bc}  \cos{((3 \theta_{bc} - 2 \pi)n + c  \theta_{bc})}  $\\
   $ r_{bc} \sin{((3 \theta_{bc} - 2 \pi)n + c  \theta_{bc})} $\\
   $ 3 h_{bc} (n + c/3)  $
  \end{tabular}
  \right ],
\text{where}
  \begin{tabular}{c}
 $ r_{bc} = \frac{3\sqrt{3}}{10} $\\
 $ h_{bc} = 1/\sqrt{10} $ \\
 $ \theta_{bc} = \arccos(-2/3) $ \text{ .}\\
  \end{tabular}      
\end{equation}

In this formula, integral values of $n$ may be taken as a node number for one rail and used to compute
its Cartesian
coordinates. Allowing $n$ to take non-integer values defines a continuous
helix in space which is close to the segmented polyline of the outer
tetrahedra edges, and equals them at integer
values.

@@

\Cref{fig:helixcomparison} illustrates this difference with a 7-tetrahedra BC helix, which is
in fact the same geometry as the robot illustrated in \cref{fig:tetrobot}.
Although the nodes coincide,
\cref{graycoxeter} evaluated
at real values generates the black helix which runs through every node, and \cref{eq:colored} defines
the red, yellow, and blue helices. (In this figure
these rail helices have been rendered at a slightly higher radius than the nodes for clarity;
in actuality the maximum distance between the continuous, curved helix and the
straight edges between nodes is much smaller than can be clearly rendered.)

The quantity $ (3 \theta_{bc} - 2 \pi) \approx 35.43 \degree $  is the angular shift between
$ \vec{V}(3n + c)= \vec{H}_{BCcolored}(n,c)$ and
$ \vec{V}(3(n+1)+c)= \vec{H}_{BCcolored}(n+1,c)$.
This quantity appears so often that we call it the ``rail angle $\rho$''. For the BC helix, $\rho_{bc} = (3 \theta_{bc} - 2 \pi)$.

@@

Note in \cref{railanglefig} the $z$-axis travel for one rail edge is denoted by $d$. In \eqref{graycoxeter} and \eqref{eq:colored}, the variable
$h$ is used for one third of the distance we name $d$. We will later justify that $d = 3h$.
In this paper we assume the length of a rail
is always $1$ as a simplification, except in proofs concenring rail length.
(We make the rail length a parameter in our JavaScript code in
\url{https://github.com/PubInv/tetrahelix/blob/master/js/tetrahelix_math.js} \cite{readtetrahelix}.)

The $\vec{H}_{BCcolored}(n,c)$ formulation can be further clarified by rewriting directly in terms of the rail angle $\rho_{bc}$ rather than $\theta_{bc}$.
Intuitively we seek an expression where $c/3$ is multiplied by a $1/3$ rotation plus the rail angle $\rho$.
We expand 
the expressions $\theta_{bc}$ and $\rho_{bc}$ in \eqref{eq:colored} and seek to isolate the term $c 2\pi/3 $.

@@

Thus, starting with the expression:
\begin{align*}
  c \theta_{bc} 
\end{align*}
we introduce 3 into the denominator...
\begin{align*}
  (c/3)  (3 \theta_{bc}) 
\end{align*}
we want $2\pi$ in numerator, so add canceling terms...
\begin{align*}
  (c/ 3) ((3 \theta_{bc} - 2 \pi)  + 2 \pi) 
\end{align*}
...and then use the definition of $\rho_{bc}$
\begin{align*}
  (c / 3) \rho_{bc}  + c 2 \pi /3 
\end{align*}
finally using algebra, we obtain
\begin{align*}
  c  ( \rho_{bc} +  2 \pi) /3  \text{.}
\end{align*}

@@

%% Actually, I am not sure if this is a clearer formulation...it might be, and need to put it in general!!!
\begin{equation}
\vec{H}_{BCcolored}(n,c) =
\left [
  \begin{tabular}{c}
    $ r  \cos{\rho_{bc} n + c (\rho_{bc} +  2 \pi) /3} $\\
   $ r  \sin{\rho_{bc} n + c (\rho_{bc} +  2 \pi) /3} $\\
   $ (n+c/3)  h_{bc} $
  \end{tabular}
  \right ],
\text{where}
  \begin{tabular}{c}
%%  $\kappa = n + c/3$ \\
    $\rho_{bc} = (3 \theta_{bc} - 2 \pi)$ \\
    $ h_{bc} = 1/\sqrt{10} $. \\    
  \end{tabular}      
\end{equation}

@@

The \emph{pitch} of any tetrahelix, defined as the axial length of a complete revolution
where $\rho \neq 0$ is:
\begin{equation}
  \label{pitcheqn}
p(\rho) = \frac{2 \pi  \cdot d}{\rho} \text{ .}
\end{equation}

The pitch of the Boerdijk-Coxeter helix of edge length $1$ is the length of three tetrahedra times this number:
\begin{align*}
   \frac{3 h_{bc}\cdot  2 \pi }{\rho_{bc}} 
  &= \frac{6 \pi}{\sqrt{10}\rho_{bc}} 
 \text{ ,}
\end{align*}
which is approximately $9.64$.

@@

A \emph{triple helix} is three congruent helices that share an axis. We show that
optimal tetrahelices are in fact triple helices with the same radius, so that all
vertices are on a cylinder. In stages, we demonstrate that optimal tetrahelices:
\begin{enumerate}
\item have the same pitch,
\item have parallel axes,
\item share the same axis,
\item have the same radius,
  \item have the same rail lengths,
  \item have axially equidistant nodes, and therefore
  \item are in fact triple helices.
\end{enumerate}

Suppose that all three rails do not have the same pitch. If we start at any shortest edge between
two rails, as we move
from node to node away from our start edge the edge lengths
between rails must always lengthen without bound,
which cannot be optimal.
So we are justified in talking about the
\emph{pitch} of 
the optimal tetrahelix as the pitch of its three rail helices, even though there are
three such helices of equivalent pitch.

@@

\begin{theorem}
  Any optimal tetrahelix with a rail angle of magnitude less than $\pi$ has all three
  axes conincident.
  \label{thm:conincident}
\end{theorem}
\begin{proof}
  Case 1: Suppose that $\rho$ is zero.
  Each helix has zero curvature, that is, it is a straight line. These lines are eqivalent
  to some three degenerate helices with conincident axes, possibly with different radii, so long as there is a phase
  term in the defintion of the helix, as in \eqref{eq:colored}. We later show the
  radii must be eqivalent.

@@

    The set of all lengths in the tetrahelix is a superset of:
    $L = \{|| \vec{R}(n) - \vec{B}(n)||\}$, which by our choice has at least one longest or shortest
    length. (Note this is just the Euclidean distance formula written as a Euclidean norm.)
    $L = \{|| \vec{R}(n) - (\vec{B'}(n) + \vec{W}(n))||\}$ and so
    $L = \{|| (\vec{R}(n) - \vec{B'}(n)) - \vec{W}(n)||\}$.
    But $\vec{R}(n) - \vec{B'}(n)$ is a constant, so the minimax value of $L$ is improved as $||\vec{W}(n)||$
    decreases.  
    By our choice that their is a Blue-to-Red edge that is either a maximum or a minimum,
    this improves the minimax value of the total tetrahelix.

@@

In principle any three helices generated with \cref{eq:triple}
has at most nine distinct edge length classes. Each edge that connects
two rails potentially has a longer length and shorter length we denote with
a $+$ or $-$. So the classes are $\{ RR, BB, YY, RB_+, RB_-, BY_+, BY_-, RY_+, RY_- \}$.
If when projecting all vertices onto the $z$-axis (dropping the $x$ and $y$ coordinates), the interval
defined by the $z$ axis value of its endpoints contains no other vertices,
we call it a \emph{one-hop} edge, and if it does contain another vertex we
call it a \emph{two-hop} edge, as illustrated in \cref{fig:naming}.
Then there are 
3 rail edges $\{ RR, BB, YY\}$,
3 one-hop lengths $\{ RB_-, BY_-, RY_- \}$ between each pair of 3 rails,
and 3 two-hop
lengths $\{ RB_{+}, BY_+, RY_+ \}$ between each pair of 3 rails,
where the two-hop length is at least the one-hop length.
However, if we generate the three helices symmetrically with \cref{eq:triple},
many of these lengths will be the same.
In fact, it is possible that there
will be only two distinct such classes, or even one, in the purely regular BC helix.

@@

Because an optimal tetrhelix has equivalent radii and equivalent pitch for all three helices,
it has equivalent rail edge lengths. Likewise, there is a single rail angle $\rho$ that
represents the rotation of two nodes connected by a single rail edge, and it is the same
for all three rails.

@@

 We show that in fact the nodes must be distributed in even thirds
 along the $z$-axis, as in fact they are in the regular BC helix.

@@

We seek a formula to generate optimal tetrahelices that accepts a
parameter that allows us to design the tetrahelix conveniently.
Please refer back to \cref{railanglefig}.
The pitch of the helix is an obvious choice, but is not defined when the
curvature is $0$, an important special case. The radius or the axial
distance between two nodes on the same rail are possible choices, but
perhaps the clearest choice is to build formulae that takes as their
input the ``rail angle'' $\rho$. We define $\rho$ to be the angle
formed in the X,Y plane $\angle \vec{H}(0,0) O \vec{H}(0,1)$ projecting out the $z$
axis and sighting along the positive $z$ axis. In other words, $\rho$
controls how far a rail edge of a tetrahelix deviates from being
parallel with the axis, or the ``twistiness'' of the tetrahelix. We use
the parameter $\chi = 1$ to indicate a chirality of counter-clockwise,
and $\chi = -1$ for clockwise. We take our coordinate system to be right-handed.

@@

Checking the important special case of the BC helix, we find that this equation
indeed holds true, treating $d$ in this equation as $3 h_{bc}$ as defined by
Gray and Coxeter, that is, $d_{bc} = 3h_{bc}$, where they are using
$h$ for the axial height from one node to
the next of a different color, but we use $d$ to mean distance between nodes of
the  same color.

@@

Rewriting our formulation in terms of $\rho$:
\begin{align*}
\numberthis  \label{eq:general}  
\vec{H}_{general}(\chi,n,c,\rho,d_{\rho},r_{\rho}) &= 
 \left [
  \begin{tabular}{c}
   $ r_{\rho} \cos(\chi \cdot (n \rho + c(\rho +  2 \pi) /3)) $\\
   $ r_{\rho}  \sin(\chi \cdot (n \rho + c(\rho +  2 \pi) /3)) $\\
   $ d_{\rho} (n + c /3) $
  \end{tabular}
  \right ] \\
\text{where: } &
\begin{tabular}{c}
  $   1 = d_{\rho}^2 + 4 r_{\rho}^2 (\sin{ \rho / 2})^2 $ \\
    $\chi \in \{-1,1\}$ \text{ .} \\  
\end{tabular}
\end{align*}

$\vec{H}_{general}$ forces the user to select three values: $\rho$, $r_{\rho}$, and $d_{\rho}$ satisfying \eqref{railangle}.

@@

\begin{proof}
By \cref{eventhirds}, we can compute the (at most) three edge-lengths of an optimal
tetrahelix by formula universally quantified by $n$ and $c$:
\begin{align*}
  \text{rail} &= || \vec{H}_{general}(n,c,\rho,d_{\rho},r) - \vec{H}_{general}(n+1,c,\rho,d_{\rho}),r)) || = 1 \\
  \text{one-hop} &= || \vec{H}_{general}(n,c,\rho,d_{\rho},r) - \vec{H}_{general}(n,c+1,\rho,d_{\rho},r)) || \text{ and,} \\
  \text{two-hop} &= || \vec{H}_{general}(n,c,\rho,d_{\rho},r) - \vec{H}_{general}(n,c+2,\rho,d_{\rho},r)) || \text{ .}\\  
\end{align*}
This syntax just represents the Euclidean distance formula. Thus:

@@

We can now give a formula for $ d_{opt} $ computed from $\rho, r_{opt}$ via the rail angle equation \eqref{railangle}:

@@

\begin{corollary}
  The equitetrabeam with minimal maximal edge ratio is produced
  \newline by $\vec{H}_{general}$ when $ r = \sqrt{\frac{8}{27}} $.
\end{corollary}

\begin{proof}
Choosing $d = 1$ and $\rho = 0$ we use Equation \eqref{eqrhoopt} to find the radius of 
optimal minimax difference.

Substituting into \eqref{eq:general}:

\begin{align*}
  \text{one-hop} &= \sqrt{\frac{1}{9} + 3r^2}\\
\end{align*}
Then:
\begin{align*}
   1  &=  \sqrt{\frac{1}{9} + 3r^2} & \text{ solved by... }\\
   r  &= \sqrt{\frac{8}{27}} &\approx 0.54 \\
\end{align*}
\end{proof}

This radius\footnote{Another interesting but non-optimal solution is derived by setting
  (one-hop + two-hop)/2 = 1,
  occurs at $r = \sqrt{35}/4$ which produces
  three length classes of $11/12, 12/12, 13/12$.}
produces a two-hop rail length of $\frac{2}{\sqrt{3}}$.
The difference between this and $1$ is $\approx 15.47\% $. The inradius of the equitetrabeam of unit
rail length from both Equation \eqref{eq:inradius} and the fact that the inradius of
an equilateral triangle is half the circumradius is $\sqrt{\frac{8}{27}}/2$, or $\frac{\sqrt{6}}{9}$.

@@

To the extent that we value tetrabeams (that is, tetrahelices with a rail angle of $0$,
and therefore zero curvature) as mathematical or engineering objects,
we have motivated the development of $\vec{H}_{general}$ as a transformation of $\vec{V}(n)$ defined by
Equation \eqref{graycoxeter} from Gray and Coxeter. It is difficult to see how
the $\vec{V}(n)$ 
formulation could ever give rise to a continuum producing the tetrabeam,
since setting the angle in that equation to zero can produce only collinear points.

@@

We observe that Equations \eqref{eqrhoopt} and \eqref{dopt} compute $r_{opt}$ and $d_{opt}$ which
create an optimal tetrahelix for any rail angle $\rho$ between $0$, which
gives the equitetrabeam and
$\rho_{bc} \approx 35.43 \degree$, which gives the BC helix.

 Because the equitetrabeam which has a rail angle of $0$ still has
 chirality, that is, one still must decide to connect the one-hop edge to
 the clockwise or the counter-clockwise node, it is not possible to build
 a smooth continuum where $\rho$ transitions from positive to negative
 which remains optimal. One can use a negative $\rho$ in $\vec{H}_{general}$
 but it does not produce minimax optimal tetrahelices. In other words,
 untwisting a counter-clockwise tetrahelix to rail angle $0$ and then going
even further does produce a clockwise tetrahelix, but one in which the
 one-hop and two-hop lengths in the wrong places, that is, two-hop
 becomes shorter than one-hop.
 Likewise, $\rho > \rho_{bc}$ generates
 a tetrahelix, but minimax optimality is not guaranteed by $\vec{H}_{general}$.

@@

\end{proof}

>>>>

The vertices of the tetrahedra lie upon
three helices about the central axis.
The \\
Tetrobot\cite{readglussbot,TetrobotBook} uses the regularity of
this geometry to make a tentacle-like robot that can crawl like a slug
or mollusk.  These modular robotic systems use mechanical actuators which can
change their length, connected by special joints,
such as the 3D printable Song-Kwon-Kim\cite{song2003spherical} joint or the CMS joint\cite{HamlinSandersonCMS}
used in the original Tetrobot,
which allow many members to meet in a single point.
Such machines can
follow purely regular mathematical models such as the Boerdijk-Coxeter
helix or the Octet Truss\cite{richard1961synergetic}.

@@

Ideally we would have a simple formula for defining the vertices based on
any curvature or pitch we choose. It is a goal of
this paper to relate the Cartesian coordinate approach and the member-length approach to
generating a tetrahelix continuum.

@@

R.W. Gray's website\cite{graytetrahelix}, repeating a formula by Coxeter\cite{coxeter1985simplicial} in a more accessible form, gives the Cartesian coordinates $\begin{bmatrix}
           x \\
           y \\
           z
         \end{bmatrix}$
for a counter-clockwise BC Helix in a right-handed coordinate system:
\begin{equation}
  \label{graycoxeter}
\vec{V}(n) =
\left [
  \begin{tabular}{c}
   $ r_{bc} \cos{n \theta_{bc}} $\\
   $ r_{bc} \sin{n \theta_{bc}} $\\
   $ n h_{bc}  $
  \end{tabular}
  \right ],
\text{where:}
  \begin{tabular}{c}
 $ r_{bc} = \frac{3\sqrt{3}}{10} \approx 0.5196 $,\\
 $ h_{bc} = 1/\sqrt{10} \approx 0.3162 $, and \\
 $ \theta_{bc} = \arccos(-2/3) $  \text{,}\\
  \end{tabular}      
\end{equation}
and where $n$ represents each integer numbered vertex in succession on every colored rail.

The apparent rotation of a vertex on an outer-edge, that is $\vec{V}(n)$ relative from $\vec{V}(n+3)$
for any integer $n$
in \cref{graycoxeter}, is $3 \theta_{bc} - 2\pi$.

This formula defines a helix, but it is not any of the apparent helices, or \emph{rail} helices, of the
BC helix, but rather one that winds much more rapidly through all
vertices. To a designer of tetrahelices, it is more natural to think of
the three helices which are visually apparent, that is, those three
which are closely approximated by the outer edges or rails of
the BC helix. We think of each of these three rails as being a different color: red, blue, or yellow.
This situation is illustrated in \cref{fig:helixcomparison}, wherein the black helix represents that
generated by Equaiton \cref{graycoxeter},
and the colored helices are generated by \cref{eq:colored}.

In order to develop the continuum of slightly irregular tetrahelices described in \cref{sec:continuum},
we need a formula that gives us the vertices of just
one rail helix, denoted by color $c$ and integer vertex number $n$:
\[
(\forall n \in \mathbb{Z}, \forall c \in \{0,1,2\} : \vec{H}_{BCcolored}(n,c) = \vec{V}(3n +c)) .
\]

Such a helix can be written:
\begin{equation}
  \label{eq:colored}
 \vec{H}_{BCcolored}(n,c) =
\left [
  \begin{tabular}{c}
   $ r_{bc}  \cos{((3 \theta_{bc} - 2 \pi)n + c  \theta_{bc})}  $\\
   $ r_{bc} \sin{((3 \theta_{bc} - 2 \pi)n + c  \theta_{bc})} $\\
   $ 3 h_{bc} (n + c/3)  $
  \end{tabular}
  \right ],
\text{where}
  \begin{tabular}{c}
 $ r_{bc} = \frac{3\sqrt{3}}{10} $,\\
 $ h_{bc} = 1/\sqrt{10} $, and \\
 $ \theta_{bc} = \arccos(-2/3) $ \text{.}\\
  \end{tabular}      
\end{equation}

In this formula, integral values of $n$ may be taken as a vertex number for one rail and used to compute
its Cartesian
coordinates. Allowing $n$ to take non-integer values defines a continuous
helix in space which is close to the segmented polyline of the outer
tetrahedra edges, and equals them at integer
values.

@@

\Cref{fig:helixcomparison} illustrates this difference with a 7-tetrahedra BC helix, which is
in fact the same geometry as the robot illustrated in \cref{fig:tetrobot}.
Although the vertices coincide,
 \cref{graycoxeter} evaluated
at real values generates the black helix which runs through every vertex, and \cref{eq:colored} defines
the red, yellow, and blue helices. (In this figure
these rail helices have been rendered at a slightly higher radius than the vertices for clarity;
in actuality the maximum distance between the continuous, curved helix and the
straight edges between vertices is much smaller than can be clearly rendered.)

The quantity $ (3 \theta_{bc} - 2 \pi) \approx 35.43 \degree $  is the angular shift between
adjacent vertices on the same rail:
$ \vec{V}(3n + c)= \vec{H}_{BCcolored}(n,c)$ and
$ \vec{V}(3(n+1)+c)= \vec{H}_{BCcolored}(n+1,c)$.
This quantity appears so often that we call it the ``rail angle $\rho$''. For the BC helix, $\rho_{bc} = (3 \theta_{bc} - 2 \pi)$.

@@

Note in \cref{railanglefig} the $z$-axis travel for one rail edge is denoted by $d$. In Equaiton \cref{graycoxeter} and  \cref{eq:colored}, the variable
$h$ is used for one third of the distance we name $d$. We will later justify that $d = 3h$.
In this paper we assume the length of a rail
is always $1$ as a simplification, except in proofs concenring rail length.
(We make the rail length a parameter in our JavaScript code in
\url{https://github.com/PubInv/tetrahelix/blob/master/js/tetrahelix_math.js} \cite{readtetrahelix}.)

The $\vec{H}_{BCcolored}(n,c)$ formulation can be further clarified by rewriting directly in terms of the rail angle $\rho_{bc}$ rather than $\theta_{bc}$.
Intuitively we seek an expression where $c/3$ is multiplied by a $1/3$ rotation plus the rail angle $\rho$.
We expand 
the expressions $\theta_{bc}$ and $\rho_{bc}$ in \cref{eq:colored} and seek to isolate the term $c 2\pi/3 $.

@@

Thus, starting with the expression:
\begin{align*}
  c \theta_{bc} 
\end{align*}
we introduce 3 into the denominator...
\begin{align*}
  (c/3)  (3 \theta_{bc}) 
\end{align*}
we want $2\pi$ in numerator, so add canceling terms...
\begin{align*}
  (c/ 3) ((3 \theta_{bc} - 2 \pi)  + 2 \pi) 
\end{align*}
...and then use the definition of $\rho_{bc}$
\begin{align*}
  (c / 3) \rho_{bc}  + c 2 \pi /3 
\end{align*}
finally we obtain:
\begin{align*}
  c  ( \rho_{bc} +  2 \pi) /3  \text{.}
\end{align*}

@@

%% Actually, I am not sure if this is a clearer formulation...it might be, and need to put it in general!!!
\begin{equation}
\vec{H}_{BCcolored}(n,c) =
\left [
  \begin{tabular}{c}
    $ r  \cos{\rho_{bc} n + c (\rho_{bc} +  2 \pi) /3} $\\
   $ r  \sin{\rho_{bc} n + c (\rho_{bc} +  2 \pi) /3} $\\
   $ (n+c/3)  h_{bc} $
  \end{tabular}
  \right ],
\text{where}
  \begin{tabular}{c}
%%  $\kappa = n + c/3$ \\
    $\rho_{bc} = (3 \theta_{bc} - 2 \pi)$, and \\
    $ h_{bc} = 1/\sqrt{10} $. \\    
  \end{tabular}      
\end{equation}

@@

The \emph{pitch} of any tetrahelix, defined as the axial length of a complete revolution
where $\rho \neq 0$ is:
\begin{equation}
  \label{pitcheqn}
p(\rho) = \frac{2 \pi  d}{\rho} \text{.}
\end{equation}

The pitch of the Boerdijk-Coxeter helix of edge length $1$ is the length of three tetrahedra times
$p(\rho_{bc})$:
\begin{align*}
   \frac{3 h_{bc}  2 \pi }{\rho_{bc}} 
   = \frac{6 \pi}{\sqrt{10}\rho_{bc}}
   \approx 9.64
 \text{.}
\end{align*}

@@

A \emph{triple helix} is three congruent helices that share an axis. We show that
optimal tetrahelices are in fact triple helices with the same radius, so that all
vertices are on a cylinder. In stages, we demonstrate that optimal tetrahelices:
\begin{enumerate}
\item have the same pitch,
\item have parallel axes,
\item share the same axis,
\item have the same radius,
  \item have the same rail lengths,
  \item have axially equidistant vertices, and therefore
  \item are in fact triple helices.
\end{enumerate}

Suppose that all three rails do not have the same pitch. If we start at any shortest edge between
two rails, as we move
from vertex to vertex away from our start edge the edge lengths
between rails must always lengthen without bound,
which cannot be optimal.
So we are justified in talking about the
\emph{pitch} of 
the optimal tetrahelix as the pitch of its three rail helices, even though there are
three such helices of equivalent pitch.

@@

\begin{theorem}
  Any optimal tetrahelix with a rail angle of magnitude less than $\pi$ has all three
  axes conincident.
  \label{thm:conincident}
\end{theorem}
\begin{proof}
  Case 1: Suppose that $\rho$ is zero.
  Each helix has zero curvature, that is, it is a straight line. These lines are eqivalent
  to some three degenerate helices with conincident axes, possibly with different radii, so long as there is a phase
  term in the defintion of the helix, as in \cref{eq:colored}. We later show the
  radii must be eqivalent.

@@

    The set of all lengths in the tetrahelix is a superset of:
    $L = \{|| \vec{R}(n) - \vec{B}(n)||\}$, which by our choice has at least one longest or shortest
    length.
    $L = \{|| \vec{R}(n) - (\vec{B'}(n) + \vec{W}(n))||\}$ and so
    $L = \{|| (\vec{R}(n) - \vec{B'}(n)) - \vec{W}(n)||\}$.
    But $\vec{R}(n) - \vec{B'}(n)$ is a constant, so the minimax value of $L$ is improved as $||\vec{W}(n)||$
    decreases.  
    By our choice that their is a Blue-to-Red edge that is either a maximum or a minimum,
    this improves the minimax value of the total tetrahelix.

@@

In principle any three helices generated with \cref{eq:triple}
has at most nine distinct edge length classes. Each edge that connects
two rails potentially has a longer length and shorter length we denote with
a $+$ or $-$. So the classes are $\{ RR, BB, YY, RB_+, RB_-, BY_+, BY_-, RY_+, RY_- \}$.
If when projecting all vertices onto the $z$-axis (dropping the $x$ and $y$ coordinates), the interval
defined by the $z$ axis value of its endpoints contains no other vertices,
we call it a \emph{one-hop} edge, and if it does contain another vertex we
call it a \emph{two-hop} edge, as illustrated in \cref{fig:naming}.
Then there are 
3 rail edges $\{ RR, BB, YY\}$,
3 one-hop lengths $\{ RB_-, BY_-, RY_- \}$ between each pair of 3 rails,
and 3 two-hop
lengths $\{ RB_{+}, BY_+, RY_+ \}$ between each pair of 3 rails,
where the two-hop length is at least the one-hop length.
However, if we generate the three helices symmetrically with \cref{eq:triple},
many of these lengths will be the same.
In fact, it is possible that there
will be only two distinct such classes. In the purely regular BC helix there is only one length.

@@

Because an optimal tetrhelix has equivalent radii and equivalent pitch for all three helices,
it has equivalent rail edge lengths. Likewise, there is a single rail angle $\rho$ that
represents the rotation of two vertices connected by a single rail edge, and it is the same
for all three rails.

@@

 We show that in fact the vertices must be distributed in even thirds
 along the $z$-axis, as in fact they are in the regular BC helix.

@@

We seek a formula to generate optimal tetrahelices that accepts a
parameter that allows us to design the tetrahelix conveniently.
Please refer back to \cref{railanglefig}.
The pitch of the helix is an obvious choice, but is not defined when the
curvature is $0$, an important special case. The radius or the axial
distance between two vertices on the same rail are possible choices, but
perhaps the clearest choice is to build formulae that takes as their
input the ``rail angle'' $\rho$. We define $\rho$ to be the angle
formed in the X,Y plane with origin $O$: $\angle \vec{H}(0,0) O \vec{H}(0,1)$ projecting out the $z$
axis and sighting along the positive $z$ axis. In other words, $\rho$
controls how far a rail edge of a tetrahelix deviates from being
parallel with the axis, or the ``twistiness'' of the tetrahelix. We use
the parameter $\chi = 1$ to indicate a chirality of counter-clockwise,
and $\chi = -1$ for clockwise. We take our coordinate system to be right-handed.

@@

Checking the important special case of the BC helix, we find that this equation
indeed holds true, treating $d$ in this equation as $3 h_{bc}$ as defined by
Gray and Coxeter, that is, $d_{bc} = 3h_{bc}$, where they are using
$h$ for the axial height from one vertex to
the next of a different color, but we use $d$ to mean distance between vertices of
the  same color.

@@

Rewriting our formulation in terms of $\rho$:
\begin{align*}
\numberthis  \label{eq:general}  
\vec{H}_{general}(\chi,n,c,\rho,d_{\rho},r_{\rho}) &= 
 \left [
  \begin{tabular}{c}
   $ r_{\rho} \cos(\chi \cdot (n \rho + c(\rho +  2 \pi) /3)) $\\
   $ r_{\rho}  \sin(\chi \cdot (n \rho + c(\rho +  2 \pi) /3)) $\\
   $ d_{\rho} (n + c /3) $
  \end{tabular}
  \right ] \\
\text{where: } &
\begin{tabular}{c}
  $   1 = d_{\rho}^2 + 4 r_{\rho}^2 (\sin{ \rho / 2})^2 $, and \\
    $\chi \in \{-1,1\}$ \text{.} \\  
\end{tabular}
\end{align*}

$\vec{H}_{general}$ forces the user to select three values: $\rho$, $r_{\rho}$, and $d_{\rho}$ satisfying \cref{railangle}.

@@

\begin{proof}
By \cref{eventhirds}, we can compute the (at most) three edge-lengths of an optimal
tetrahelix by formula universally quantified by $n$ and $c$:
\begin{align*}
  \text{rail} &= || \vec{H}_{general}(n,c,\rho,d_{\rho},r) - \vec{H}_{general}(n+1,c,\rho,d_{\rho}),r)) || = 1 \text{ ,} \\
  \text{one-hop} &= || \vec{H}_{general}(n,c,\rho,d_{\rho},r) - \vec{H}_{general}(n,c+1,\rho,d_{\rho},r)) || \text{ and,} \\
  \text{two-hop} &= || \vec{H}_{general}(n,c,\rho,d_{\rho},r) - \vec{H}_{general}(n,c+2,\rho,d_{\rho},r)) || \text{ .}\\  
\end{align*}
This syntax just represents the Euclidean distance between vertices. Thus:

@@

We can now give a formula for $ d_{opt} $ computed from $\rho, r_{opt}$ via the rail angle equation \cref{railangle}:

@@

\begin{corollary}
  The equitetrabeam with minimal maximal edge ratio is produced
  \newline by $\vec{H}_{general}$ when $ r = \sqrt{\frac{8}{27}} $ and the longest member is $\frac{2}{\sqrt{3}}$,
  which is less than $16\%$ longer
  than the shortest member.
\end{corollary}

\begin{proof}
Choosing $d = 1$ and $\rho = 0$ we use  \cref{eqrhoopt} to find the radius of 
optimal minimax difference. Substituting into  \cref{eq:general} and setting one-hop\footnote{Before developing the optimal solution  \cref{eqrhoopt} we found another interesting but non-optimal tetrabeam of zero curvature
by setting
the mean of one-hop and two-hop to 1 ($(\text{one-hop} + \text{two-hop})/2 = 1$),
  which gives $r = \sqrt{35}/4$ and
  three length classes of $\frac{11}{12}, \frac{12}{12}, \frac{13}{12}$.}
 to $1$
gives:
%% \begin{align*}
%%   1  &=  \sqrt{\frac{1}{9} + 3r^2} \\
%% \end{align*}  
%% and solving for $r$, gives
%% \begin{align*}
%%    r  &= \sqrt{\frac{8}{27}} \approx 0.54 \text{.}\\
%% \end{align*}
$ 1  =  \sqrt{\frac{1}{9} + 3r^2} $, or $ r  = \sqrt{\frac{8}{27}}$. This radius
produces a two-hop edge length of $\frac{2}{\sqrt{3}} \approx 1.1547$.
\end{proof}

The inradius of the equitetrabeam of unit
rail length from both  \cref{eq:inradius} and the fact that the inradius of
an equilateral triangle is half the circumradius is $\sqrt{\frac{8}{27}}/2$, or $\frac{\sqrt{6}}{9}$.

@@

To the extent that we value tetrabeams (that is, tetrahelices with a rail angle of $0$,
and therefore zero curvature) as mathematical or engineering objects,
we have motivated the development of $\vec{H}_{general}$ as a transformation of $\vec{V}(n)$ defined by
 \cref{graycoxeter} from Gray and Coxeter. It is difficult to see how
the $\vec{V}(n)$ 
formulation could ever give rise to a continuum producing the tetrabeam,
since setting the angle in that equation to zero can produce only collinear points.

@@

We observe that  \cref{eqrhoopt,dopt} compute $r_{opt}$ and $d_{opt}$ which
create an optimal tetrahelix for any rail angle $\rho$ between $0$, which
gives the equitetrabeam and
$\rho_{bc} \approx 35.43 \degree$, which gives the BC helix.

 Because the equitetrabeam which has a rail angle of $0$ still has
 chirality, that is, one still must decide to connect the one-hop edge to
 the clockwise or the counter-clockwise vertex, it is not possible to build
 a smooth continuum where $\rho$ transitions from positive to negative
 which remains optimal. One can use a negative $\rho$ in $\vec{H}_{general}$
 but it does not produce minimax optimal tetrahelices. In other words,
 untwisting a counter-clockwise tetrahelix to rail angle $0$ and then going
even further does produce a clockwise tetrahelix, but one in which the
 one-hop and two-hop lengths in the wrong places, that is, two-hop
 becomes shorter than one-hop.
 Likewise, $\rho > \rho_{bc}$ generates
 a tetrahelix, but minimax optimality is not guaranteed by $\vec{H}_{general}$.

@@
